---
title: "Analyzing World Economic, Social and Demographic Trends"
author: "Grant Cook"
date: "4/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The website [gapminder.org](gapminder.org) visualizes social, economic and demographic data by country to debunk common misconceptions in global data trends. Hans Rosling, one of the organization's founders, gave a [famous TED Talk](https://www.youtube.com/watch?v=jbkSRLYSojo&t=59s) in which he animates various charts to show how these indicators evolve over time. I will attempt to recreate some of those visualizations here using data sourced from Gapminder and including variables on GPD and life expectancy by country over time.


```{r}
# Import libraries and dataset
library(tidyverse)
library(janitor)
library(wesanderson)

data <- read_csv("data-processed/country_5yr.csv")
```

```{r}
# Basic data summarizing
data %>%
  filter(year %in% c(2010)) %>% 
  group_by(region, year) %>% 
  summarize(Avg_GDP_pc=(sum(gdp, na.rm=TRUE)/sum(population, na.rm=TRUE)),
            Avg_LE=mean(life_expectancy, na.rm=TRUE))
```

```{r}
# Basic visualization: scatterplot of life exp vs. gdp pc in 2010
pal <- c("#009E73", "#F0E442", "#0072B2", "#D55E00")
data %>% 
  filter(year==2010) %>% 
  ggplot(aes(x=gdp_percap, y=life_expectancy, size=population, color=region)) +
  geom_point(alpha=0.8) +
  scale_color_manual(values=pal) +
  scale_x_continuous(trans="log2",
                     breaks=c(0,1000,2000,4000,8000,16000,32000,64000,128000)) +
  scale_size_continuous(range=c(1,12))

```

